A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2020-09-30 cusp_j/paper/paper.tex
pouet

\definecolor{darkgreen}{HTML}{009900}
\usepackage[normalem]{ulem}
\newcommand{\sphi}[1]{\hat{{\bf S}}_{#1}}
\newcommand{\overlap}[2]{\langle #1 | #2 \rangle}
\newcommand{\matelem}[3]{\langle #1 | #2 | #3 \rangle}
\newcommand{\deriv}[3]{\frac{\partial^{#3} #1}{\partial {#2}^{#3}}}
\newcommand{\bd}[1]{{\bf {#1}}}
\newcommand{\br}[0]{{\bf {r}}}
\newcommand{\bs}[0]{{\bf {s}}}
\newcommand{\dr}[1]{\text{d}{\bf {#1}}}
\newcommand{\psiex}[0]{\Psi^{\text{ex}}_0}
\newcommand{\phiex}[0]{\Phi^{\text{ex}}_0}
\newcommand{\phimu}[0]{\Phi^{\text{ex},\mu}_0}
\newcommand{\phimub}[0]{\Phi^{\mathcal{B},\mu}_0}
\newcommand{\psimub}[0]{\Psi^{\mathcal{B},\mu}_0}
\newcommand{\phiimub}[0]{\Phi^{\mathcal{B},\mu}_i}
\newcommand{\basis}[0]{\mathcal{B}}
\newcommand{\energyex}[0]{E^{\text{ex}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\identity}{\mathds{1}}
\newcommand{\mur}[1]{\mu({\bf r_{#1}})}

@@

\subsubsection{The eigenvectors of $\tilde{H}[\mu]$}
As shown in the previous section, the form of $\tilde{H}[\mu]$ depends on $\mu$, going from the usual repulsive hermitian Hamiltonian to a non hermitian operator which can have

\section{Detailed numerical study of $\tilde{H}[\mu]$ in the case of he He atom}
The first part of our study concerns the convergence of the ground state eigenvalue of $\tilde{H}[\mu]$ with respect to the basis set, and more precisely the dependence on $\mu$ of such convergence. 
Therefore, considering a given basis set $\basis$ and the corresponding projector $P_\basis$, one can define the transcorrelated Hamiltonian projected in a given basis set by
\begin{equation}
 \tilde{H}[\mu]^{\basis} = P_\basis \tilde{H}[\mu] P_\basis,
\end{equation}
and its right eigenvectors and eigenvalues satisfy of course
\begin{equation}
 \tilde{H}[\mu]^{\basis} \ket{\phimub} = E_i^{\basis}[\mu] \ket{\phimub}. 
\end{equation}

As pointed in section \ref{sec:h_mu_lim}, by varying $\mu$ one can go from a relatively usual repulsive Hamiltonian (even though it is non hermitian an non divergent), passing by an almost attractive Hamiltonian and even to a purely attractive Hamiltonian. 
Considering such variation of $\tilde{H}[\mu]$ when varying $\mu$,  
its projection into a basis set $\tilde{H}[\mu]^{\basis}$ will behave differently as a function of $\mu$ 
and so will $\ket{\phimub}$ and $E_i^{\basis}[\mu]$. 
Therefore, we would like to study the mathematical behaviour of $\ket{\phimub}$ and $E_i^{\basis}[\mu]$ as a function of both $\basis$ and $\mu$ to see whether the use of the new form of Jastrow factor $j(r_{12},\mu)$ can improve the convergence toward the exact ground state energy and eigenfunction. 
\subsection{$\mu$ dependence of the convergence of the total energy with respect to the basis set }
\subsubsection{Computational details}
We implemented all integrals needed for the computation of the matrix elements of $\tilde{H}[\mu]$ on a usual orbital basis set (see annexes for more details) in the case of an atomic two electron systems, together with the modification of the Slater rules due to the presence of the non hermitian term in $\tilde{H}[\mu]$. 
All implementation was done as a plugin of the program quantum package\cite{QP2}. 
Once the matrix elements of $\tilde{H}[\mu]$ are built on a given MO basis, we solve the two-body problem giving full flexibility to $\Phi_\mu$ (\textit{i.e.} it is expressed as a linear combination of all possible Slater determinants within a given Hilbert space defined with a MO basis) and use an non hermitian eigensolver to obtain both right and left eigenvectors. If not explicitly mentioned, the canonical RHF MOs are used in this work. 
The study was performed using the augmented Dunning basis aug-cc-pVXZ (X=D,T,Q,5), hereafter referred as AVXZ.

>>>>

\definecolor{darkgreen}{HTML}{009900}
\usepackage[normalem]{ulem}
\newcommand{\sphi}[1]{\hat{{\bf S}}_{#1}}
\newcommand{\overlap}[2]{\langle #1 | #2 \rangle}
\newcommand{\matelem}[3]{\langle #1 | #2 | #3 \rangle}
\newcommand{\deriv}[3]{\frac{\partial^{#3} #1}{\partial {#2}^{#3}}}
\newcommand{\bd}[1]{{\bf {#1}}}
\newcommand{\br}[0]{{\bf {r}}}
\newcommand{\bs}[0]{{\bf {s}}}
\newcommand{\dr}[1]{\text{d}{\bf {#1}}}
\newcommand{\psiex}[0]{\Psi^{\text{ex}}_0}
\newcommand{\phiex}[0]{\Phi^{\text{ex}}_0}
\newcommand{\phimu}[0]{\Phi^{\text{ex},\mu}_0}
\newcommand{\phimub}[0]{\Phi^{\mathcal{B},\mu}_0}
\newcommand{\xhimub}[0]{X^{\mathcal{B},\mu}_0}
\newcommand{\psimub}[0]{\Psi^{\mathcal{B},\mu}_0}
\newcommand{\phiimub}[0]{\Phi^{\mathcal{B},\mu}_i}
\newcommand{\basis}[0]{\mathcal{B}}
\newcommand{\energyex}[0]{E^{\text{ex}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\identity}{\mathds{1}}
\newcommand{\mur}[1]{\mu({\bf r_{#1}})}

@@

\section{Detailed numerical study of $\tilde{H}[\mu]$ in the case of he He atom}
In the present paper, we study the behaviour of $\tilde{H}[\mu]$ in the case of the isoelectronic series of the helium atom, and in that section we focus on the helium atom itself. 
The first part of our study concerns the convergence of the eigenvalues and eigenvectors of $\tilde{H}[\mu]$ with respect to the basis set, and more precisely the dependence on $\mu$ of such convergence.

Considering a given basis set $\basis$ and the corresponding projector $P_\basis$, one can define the transcorrelated Hamiltonian projected on $\basis$ by
\begin{equation}
 \tilde{H}[\mu]^{\basis} = P_\basis \tilde{H}[\mu] P_\basis,
\end{equation}
and its eigenvalues $E_i^{\basis}[\mu]$ and left/right eigenvectors satisfy of course
\begin{equation}
 \tilde{H}[\mu]^{\basis} \ket{\phimub} = E_i^{\basis}[\mu] \ket{\phimub}. 
\end{equation}
\begin{equation}
 \bigg(\tilde{H}[\mu]^{\basis}\bigg)^\dagger \ket{\xhimub} = E_i^{\basis}[\mu] \ket{\xhimub}. 
\end{equation}

As pointed in section \ref{sec:h_mu_lim}, even though $\tilde{H}[\mu]$ is still non hermitian as long as $\mu < \infty$, by varying $\mu$ one can change the nature of the electron-electron interaction in $\tilde{\mathcal{W}}_{ee}(r_{12},\mu)$ from a repulsive and non divergent interaction at large $\mu$ (an non divergent), passing by an almost attractive Hamiltonian and even to a purely attractive Hamiltonian. 
Considering such variation of $\tilde{H}[\mu]$ when varying $\mu$,  
its projection into a basis set $\tilde{H}[\mu]^{\basis}$ will behave differently as a function of $\mu$ 
and so will $\ket{\phimub}$ and $E_i^{\basis}[\mu]$. 
Therefore, we would like to study the mathematical behaviour of $\ket{\phimub}$ and $E_i^{\basis}[\mu]$ as a function of both $\basis$ and $\mu$ to see whether the use of the new form of Jastrow factor $j(r_{12},\mu)$ can improve the convergence toward the exact ground state energy and eigenfunction. 
\subsection{$\mu$ dependence of the convergence of the total energy with respect to the basis set }
\subsubsection{Computational details}
We implemented all integrals needed for the computation of the matrix elements of $\tilde{H}[\mu]$ on a usual orbital basis set (see annexes for more details) in the case of an atomic two electron systems, together with the modification of the Slater rules due to the presence of the non hermitian term in $\tilde{H}[\mu]$. 
All implementation was done as a plugin of the program quantum package\cite{QP2}. 
Once the matrix elements of $\tilde{H}[\mu]$ are built on a given MO basis, we solve the two-body problem giving full flexibility to $\Phi_\mu$ (\textit{i.e.} it is expressed as a linear combination of all possible Slater determinants within a given Hilbert space defined with a MO basis) and use an non hermitian eigensolver to obtain both right and left eigenvectors. If not explicitly mentioned, the canonical RHF MOs are used in this work. 
The study was performed using the augmented Dunning basis aug-cc-pVXZ (X=D,T,Q,5), hereafter referred as AVXZ.
